Funding programmes for early career researchers influence their independence

### 2023-09-30 article.Rmd
updated introduction

---
title: "Funding programmes for early career researchers influence their independence"
author: "David Janků"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  word_document: default
always_allow_html: yes
bibliography: researcher_independence.bib
csl: apa.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(targets)
library(tidyr)
library(dplyr)
library(cobalt)
library(knitr)
library(psych)
library(kableExtra)
library(ggplot2)
library(lme4)
library(lmerTest)
library(MuMIn)
library(ez)
library(plm)
library(effectsize)
library(nlme)

@@

@@

@@

Since the explicit goal of many funding instruments that are being used by research funders all over the world is to help early career researchers become independent, we would like to use the newly created indicators to assess the extent, to which such programmes influence researchers independence.

H1: Researchers, who received funding from programmes aimed at funding early career researchers, will show more independence on their PhD advisor 5 years after receiving the funding than researchers who did not receive this kind of funding.

@@

For the matched researchers from control groups (see the matching description below in the Analysis section), we didn't search for their supervisors manually, but rather used the "most frequent co-author of the first 5 publications" as a proxy to identify supervisors. Crosschecking this proxy with the actual data about supervisors we have, we found that this algorithm was able to identify supervisors correctly in about 50 % of cases (see the Attachment 1). Using this method, we were able to find `r (nrow(tar_read("final_data") %>% filter(treatment != 1) %>% filter(!is.na(sup_name))) / nrow(tar_read("matching") %>% filter(treatment != 1)))*100` % of the observations in the matched control groups.

@@

@@

>>>>

---
title: "Funding programmes for early career researchers influence their independence"
author: "David Janků"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  word_document: default
always_allow_html: yes
bibliography: merged.bib
csl: apa.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
library(targets)
library(tidyr)
library(dplyr)
library(cobalt)
library(knitr)
library(psych)
library(kableExtra)
library(ggplot2)
library(lme4)
library(lmerTest)
library(MuMIn)
library(ez)
library(plm)
library(effectsize)
library(nlme)

@@

## Intro 1: focus on the bigger picture, sociology and policy perspective

@@

Early career researcher independence is central to the long-lasting debate about how science progresses - whether it is indeed guided mostly by the truth and empirical evidence, or whether more social and demographic factors might play an important role. @hull1978 found that even more senior researchers readily updated their beliefs with the new evidence, supporting the former, though many (famously including Max Planck) observed that senior researchers were more fixed in their beliefs and new ideas were mostly brought up by the incoming early career researchers giving rise to the “science moves forward one funeral at the time” proverb. More recently, @azoulay2010 provided further support for the latter by investigating the effects of unexpected deaths of prominent scientists on their research fields and their collaborators' careers. They found a decrease in the productivity of junior collaborators and an increase in published work by young newcomers to the field rewarded by an increased citation rate. That suggests the sudden deaths of prominent scientists gave room for more disruptive changes in the field and for bringing new ideas and perspectives while harming the status quo holders who worked with the prominent scientists on their research agendas - a “Goliath effect”, as coined by @wang2021, p.42).

/// here I can potentially talk a bit more about Inheritance of research agendas - how often does that happen, etc..///

Insofar as the latter is true, in order to accelerate scientific progress, we might want to tackle this trend by supporting early career researchers to gain more independence from their PhD advisors. Younger researchers are already more likely to come up with new topics and ideas, as found by @packalen2019 by looking at the usage of new words in research papers. However, they might still feel various kinds of pressures to stick with their mentor’s research agendas and lack of opportunities to become more independent. In recent decades, there has been a worry that researchers’ independence is decreasing, illustrated by an increase in age at which US biomedical researchers received their first grant (NAS, 2005; @danielsGenerationRiskYoung2015), as well as direct measures of new researchers becoming more dependent on their PhD advisors across almost all disciplines in Slovenia (@rojkoScientificPerformanceResearch2022, table 6 on p.9).

Perhaps to tackle this trend, many grant agencies decided to start awarding grants for early career researchers, usually up to 8 years after finishing their PhD. Such programmes have often explicitly stated goals of helping researchers become more independent, set up their own research group and open up new lines of research. The examples of such programmes could be Starting Grants by European Research Council (@StartingGrantERC), Junior project and Junior Star by Czech Science Foundation (2020), Walter N. Benjamin Programme by DFG (@WalterBenjaminProgramme), Innovation Research Incentives Scheme by Netherlands Organization of Scientific Research (NWO) (2021).

Moreover, stimulating independence and supporting researchers to open up new lines of research is also a worthy policy goal regardless of researchers’ age. @rzhetsky2015 found that biomedical researchers are in practice often choosing less exploratory strategies for topic choice than would be socially optimal if we wanted to explore the cognitive map of science (or in this case chemicals) as fast as possible, and their choices become more conservative over time. They suggest that increased interdisciplinarity, risk-taking and the publication of experimental failures would substantially improve the speed of discovery.

//// factors influencing independence////
There are few known factors influencing the level of research independence. One of them is study discipline. Based on some early empirical studies (@rojkoScientificPerformanceResearch2022), it seems that researchers in disciplines such as philosophy, psychology; religion, theology; social sciences; the arts, recreation, entertainment, sport; language, linguistics, literature; and geography, biography, history experience significantly higher levels of independence on their PhD advisor than researchers in disciplines such as science and knowledge, organization, computer science, information, documentation, librarianship, institutions, publications; mathematics, natural sciences; and applied sciences, medicine, technology.

Interestingly, there is also some evidence that the propensity to cognitive independence might be inherited and passed down from generation to generation in mentor-mentee relationships @yoshioka-kobayashi2021 - perhaps it is not only research agendas that students inherit from their advisors but also the tendency to deviate from those research agendas. This might also suggest that to increase junior researchers’ independence, we might want to teach their advisors to be more independent, in order to have a ripple effect down the line.

Finally, there is also some preliminary evidence that higher independence modestly correlates with the number of citations acquired (@rojkoScientificPerformanceResearch2022), but others argue that even the research systems with very low independence can create a lot of scientific impact, and thus the relationships between independence and impact might not be as clear (@glaserIndependenceResearchReview2022)

Although researcher independence seems to be an important feature that is acknowledged by many actors in the research systems and that research funders try to stimulate, there hasn’t been much effort focused on quantifying this phenomenon. Some studies try to measure independence in a simple way, as a ratio of publications the given researcher produced without their supervisor divided by the total number of publications of the given researchers (e.g. @rojkoScientificPerformanceResearch2022). However, such an indicator could be misleading in the world of team science, and also only focuses on collaborations, not tracking whether a given researcher becomes independent in their thinking and research topics. [@vandenbesselaarMeasuringResearcherIndependence2019] therefore proposed a more complex indicator for measuring researcher independence on the individual level, based on co-authorship network and topical overlaps. Their measure uses only information about publications’ characteristics (authorship and reference list) regardless of their impacts, unlike other alternative measures that focus solely on the impact of the independent work in terms of received citations (Dey, et al., 2021, @ALISNovelMetric).

Since the explicit goal of many funding instruments that are being used by research funders all over the world is to help early career researchers become independent, we decided to use the newly created indicators to assess the extent, to which such programmes influence researchers' independence.

## Intro 2: Perspective of “What are the effects of competitive grant funding? And looking at independence as one of such effects”

///Primer of grant-making context:

Grant funding mechanisms have previously been shown to stimulate productivity in terms of the number of publications, impact in terms of citations, and even scientific novelty (@wang2018, or as some authors argue rather interdisciplinarity: @fontana2020). In terms of researchers’ careers, obtaining grant funding was connected to significantly higher chances of career progression and obtaining a professorship (@bloch2014).

///Though there is still some disagreement about the direction and magnitude of some of these effects, c.f. @@glaserIndependenceResearchReview2022 on novelty. --> here I can potentially go more in-depth into the critiques of grant funding mechanisms? E.g. spell out the critiques, such that they are strengthening the Matthews effect, they reduce novelty, etc....

However, there are still several components of research excellence (see @vanarensbergenDifferentViewsScholarly2014; @moherAssessingScientistsHiring2018) that have not been investigated yet, with one having specifically important effects on researchers’ carers and development: their independence. Researchers’ independence from their PhD advisor is one of the criteria often used for assessing researchers for hiring, tenure and promotion (@vanarensbergenDifferentViewsScholarly2014; @moherAssessingScientistsHiring2018). The promise of independence is that researchers will “challenge the existing common views and open up new lines of research” (@vandenbesselaarMeasuringResearcherIndependence2019, p.2).

@@

Since the explicit goal of many funding instruments that are being used by research funders all over the world is to help early career researchers become independent, we used the newly created indicators to assess the extent, to which such programmes influence researchers independence.

H1: Researchers, who received funding from programmes aimed at funding early career researchers, will show more independence from their PhD advisors after receiving the funding than researchers who did not receive funding.

H2: Researchers, who received funding from programmes aimed at funding early career researchers, will show more independence from their PhD advisors compared to similar researchers who received other types of funding at similar point in theri career.

## Intro 3: focusing on support for early career researchers (and why is that important topic) and then mention the independence as one of the ways how we can test whether our support mechanisms for early stage researchers actually work

Funding for early career researchers is an important topic in science funding policy discussions as it might have a significant impact on career progression of the new generations of researchers (@wright2017; @pickett2019; @vandenbesselaar2015). At the same time, the age of receiving a first grant has been growing significantly over the past decades (@powell2016, @christian2021). These trends have motivated some to create new proposals for funding agencies and other research funders in order to support early career researchers (@dewinde2021). Some funding agencies decided to tackle this trend by giving advantage to first-awardees and early career researchers in their usual distribution process, while others have created specific programmes designated to funding young researchers soon after their PhD.

///maybe talk about other types of support that early career researchers get? Or maybe have another paragraph describing what does the life of early career researchers usually look like? (--> perhaps mentioning that the time to finish PhD gets longer and longer, people are now expected to do multiple post-docs, etc... --> perhaps borrow some info from Science of Science book Careers section)///

It is expected, that by giving early-stage researchers funding, they will be able to develop more independence and open up new lines of research (@vandenbesselaarMeasuringResearcherIndependence2019, p.2). Younger researchers are already more likely to come up with new topics and ideas, as found by [@packalen2019] by looking at the usage of new words in research papers. However, they might still feel various kinds of pressures to stick with their mentor’s research agendas and lack of opportunities to become more independent. In recent decades, there has been a worry that researchers’ independence is decreasing, illustrated by increase in age at which US (biomedical) researchers received their first grant (NAS, 2005; @danielsGenerationRiskYoung2015), as well as direct measures of new researchers becoming more dependent on their PhD advisors across almost all disciplines in Slovenia (@rojkoScientificPerformanceResearch2022, table 6 on p.9).
///talk about the balance between experience and independence??///

Perhaps to tackle this trend, many grant agencies decided to start awarding grants for early career researchers, usually up to 8 years after finishing their PhD. Such programmes have often explicitly stated goals of helping researchers become more independent, set up their own research group and open up new lines of research. The examples of such programmes could be Starting Grants by European Research Council (@StartingGrantERC), Junior project and Junior Star by Czech Science Foundation (2020), Walter N. Benjamin Programme by DFG (@WalterBenjaminProgramme), Innovation Research Incentives Scheme by Netherlands Organization of Scientific Research (NWO) (2021)... something from the US would be also good)

Since the explicit goal of many funding instruments that are being used by research funders all over the world is to help early career researchers become independent, we would like to use the newly created indicators to assess the extent, to which such programmes influence researchers' independence.

@@

For the matched researchers from control groups see the matching description below in the Analysis section.

@@

We also tested an assumption that had a hope of reducing time intensity of data collection by substitution it for algorithmic approach: using the "most frequent co-author of the first 5 publications" as a proxy to identify supervisors. Crosschecking this proxy with the actual data about supervisors we found manually, we found that this algorithm was able to identify supervisors correctly in about XXXXX % of cases (see the Attachment 1).

@@

## construction of the independnece identificator

Correlation of its components

correlation with other stuff

- test whether results for the "cognitive" part are the same as results for the "collaboration" part -> do all the tests
